---
title: "IgXplore_IGH"
output: html_document
date: "2026-05-05"
---
```{r, global.options, include=FALSE}
# Global knitr settings 
# Set up figures and chunks messages
 
knitr::opts_knit$set(
  echo = TRUE,
  root.dir = getwd(),
  fig.width = 6, fig.height = 5,
  fig.align = "center",
  out.width = 768,
  fig.pos = "H",
  warning = FALSE,
  message = FALSE
)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.width = 6, fig.height = 5,
  fig.align = "center",
  out.width = 768,
  fig.pos = "H"
)
 

result.dir <- paste0("../results/IgXplore/", Sys.Date(), "/")
figures.dir <- paste0(result.dir, "Graphs/")
 
# Create result.dir with date if not existent
ifelse(isFALSE(dir.exists(result.dir)),
  dir.create(result.dir, recursive = TRUE),
  "Result directory for today exists already!"
)
 
# Create figures.dir with date if not existent
ifelse(isFALSE(dir.exists(figures.dir)),
  dir.create(figures.dir, recursive = TRUE),
  "Figures directory for today exists already!"
)
 
options(stringsAsFactors = FALSE)
```

Libraries
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggsci)
library(stringr)
library(circlize)
library(ggprism)
library(purrr)
library(iNEXT)
library(ggpubr)
library(forcats)
library(ggalluvial)
library(vegan)
library(ggseqlogo)
library(scifer)
library(tidytext)
library(lme4)
library(lmerTest) 
library(ggridges)
```

Setting the theme
```{r}
theme_set(
  theme_prism() +
    theme(
      aspect.ratio = 1,
      legend.position = "right"
    )
)

#colours
scale_fill_malaria <- function(...) {
  scale_fill_manual(
    values = c(
      "blood_D16" = "#92C851",
      "blood_D49" = "#C85157", 
      "blood_D66" = "#946edb",
      "BM_D66"    = "#51C8C2",
      "203hi-LMQ" = "#FF6004",
      "203lo-LMQ" = "#FFB58A",
      "203hi-MM"  = "#1C4A87",
      "203lo-MM"  = "#9EBFEB",
      "MM"        = "#1FB4FF",
      "LMQ"       = "#C82909"
    ),
    ...
  )
}

scale_color_malaria <- function(...) {
  scale_color_manual(
   values = c(
     "blood_D16" = "#92C851",
      "blood_D49" = "#C85157", 
      "blood_D66" = "#946edb",
      "BM_D66"    = "#51C8C2",
      "203hi-LMQ" = "#FF6004",
      "203lo-LMQ" = "#FFB58A",
      "203hi-MM"  = "#1C4A87",
      "203lo-MM"  = "#9EBFEB",
      "MM"        = "#1FB4FF",
      "LMQ"       = "#C82909"
    ),
    ...
  )
}
```

Data
```{r}
data <- readRDS("~/Desktop/Loré Lab/BCR/BCR_analysis/data/260303_aa_lineage_clonotypes_all_igxplore.rds")

metadata <- read.csv("../data/metadata_file.tsv", sep = "\t")
metadata$animal_ID <- as.character(metadata$animal_ID)

df <- read.table("/Users/ju-hye.lim/Desktop/Loré Lab/BCR/BCR_analysis/data/igblast_output_with_c.tsv", sep = '\t', header = TRUE)
```

Data wrangling
```{r}
data <- data %>%
  mutate(
    tissue = case_when(
      grepl("D16", timepoint) ~ "blood_D16",
      grepl("smartseq_blood", id) ~ "blood_D66",
      grepl("smartseq_BM", id) ~ "BM_D66",
      grepl("smartseq_MBC_blood", id) ~ "BM_D66",
      grepl("M4FGR", sequence_id) ~ "blood_D66",
      grepl("LW99H", sequence_id) ~ "BM_D66",
      grepl("D49", timepoint) ~ "blood_D49",
      TRUE ~ "unknown"
    )
  ) %>%
  mutate(animal_ID = case_when(
    grepl("D16", timepoint) ~ str_sub(sequence_id,1, 4),
    str_starts(id, "smartseq") ~ "5013",
    str_starts(id, "S") ~ str_extract(id, "(?<=S)\\d+"),
    str_starts(id, "M4FGR") ~ "3001",
    grepl("D49", timepoint) ~ str_sub(sequence_id,1, 4),
    TRUE ~ NA_character_
))%>%
  left_join(metadata, by = "animal_ID")

#making a cdr3_aa_length column based on cdr3_aa col
data <- data %>%
  mutate(cdr3_aa_length = nchar(cdr3_aa))

#removing alleles from v_call and j_call
data <- data %>%
  mutate(v_gene = str_extract(v_call, "^[^*]+"))

data <- data %>%
  mutate(j_gene = str_extract(j_call, "^[^*]+"))

#single cell sanger data
data_sc <- data %>%
filter(tissue %in% c("blood_D16", "blood_D49"))

#bulk data
data_bulk <- data %>%
filter(grepl("M4FGR|LW99H", sequence_id))

#removing SHM <= 20 + outlier removal
data_filtered <- data %>%
  group_by(tissue) %>%
  filter(
    V_SHM >= 0,                                          # no negative SHM
    V_SHM <= 20,                                         # biological ceiling
    V_SHM <= quantile(V_SHM, 0.99, na.rm = TRUE)        # remove top 1% per timepoint
  ) %>%
  ungroup()

summary <- data %>%
  group_by(group, tissue) %>%
  summarise(n = n())
```

Statistical comparisons
```{r}
comparisons_tissues <- list(
  c("blood_D16", "blood_D66"),
  c("blood_D16", "blood_D49"),
  c("blood_D49", "blood_D66"),
  c("blood_D49", "BM_D66"),
  c("blood_D16", "BM_D66"),
  c("BM_D66", "blood_D66")
)

comparisons_groups <- list(
  c("203hi-LMQ", "203hi-MM"),
  c("203hi-LMQ", "203lo-MM"),
  c("203hi-MM",  "203lo-MM")
)

comparisons_adj <- list(
  c("MM", "LMQ")
)
```


******************************************

somatic hypermutation (v_shm)
```{r}

# add n to x axis labels
n_label <- data_filtered %>% 
  group_by(tissue) %>% 
  summarise(n = n()) %>%
  mutate(label = paste0(tissue, "\n(n=", n, ")"))

ggplot(data_filtered, aes(x = tissue, y = V_SHM, fill = tissue)) +
  geom_boxplot() +
    scale_fill_malaria()+
  stat_compare_means(comparisons = comparisons_tissues,
                     method = "wilcox.test", 
                     label = "p.signif") +
  scale_x_discrete(labels = setNames(n_label$label, n_label$tissue))+
  labs(
       x = "Timepoint",
       y = "% of Somatic Hypermutation")+
  theme(legend.position = "none", axis.text = element_text(size = 16), axis.title = element_text(size = 18))


#save
ggsave(filename = paste0(figures.dir, "IGHV_SHM_timepoint_v3.svg"), width = 10, height = 6)


data_filtered %>%
  ggplot(aes(x = V_SHM, y = tissue, fill = tissue)) +
  geom_density_ridges(
    alpha = 0.7, scale = 2.5, bandwidth = 0.5, 
    from = 0, to = 20) +
  scale_fill_malaria() +
  scale_y_discrete(expand = expansion(mult = c(0.05, 0.3))) +
  labs(x = "% Somatic Hypermutation", y = NULL) +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 18), legend.position = "none")

#save
ggsave(filename = paste0(figures.dir, "IGHV_SHM_timepoint_dist.svg"), width = 10, height = 6)

```

******************************************
CDR3 length by aa
```{r}

n_label <- data %>% 
  group_by(tissue) %>% 
  summarise(n = n()) %>%
  mutate(label = paste0(tissue, "\n(n=", n, ")"))

ggplot(data, aes(x = tissue, y = cdr3_aa_length, fill = tissue)) +
  geom_boxplot() +
  scale_fill_malaria()+
  stat_compare_means(comparisons = comparisons_tissues,
                     method = "wilcox.test", 
                     label = "p.signif") +
  scale_x_discrete(labels = setNames(n_label$label, n_label$tissue))+
  theme(legend.position = "none", axis.text = element_text(size = 16), axis.title = element_text(size = 18)) +
  labs(x = "Timepoint",
       y = "CDR3 Length (AA)")

#save
ggsave(filename = paste0(figures.dir, "IGH_CDR3_length_timepoint.svg"), width = 10, height = 6)

#distribution
data %>%
  ggplot(aes(x = cdr3_aa_length, y = tissue, fill = tissue)) +
  geom_density_ridges(
    alpha = 0.7, scale = 2.5, bandwidth = 0.5, 
    from = 4, to = 22) +
  scale_fill_malaria() +
  scale_y_discrete(expand = expansion(mult = c(0.05, 0.3))) +
  labs(x = "CDR3 Length (AA)", y = "Timepoint") +
  theme(legend.position = "none", axis.text = element_text(size = 16), axis.title = element_text(size = 18))

#save
ggsave(filename = paste0(figures.dir, "IGH_CDR3_length_timepoint_dist.svg"), width = 10, height = 6)

```



******************************************

IGHV Gene usage (%) by timepoint (top 5 of each timepoint shown)
```{r}
data %>%
  count(tissue, v_gene) %>%
  group_by(tissue) %>%
  mutate(proportion = n / sum(n) * 100) %>%
  top_n(10, n) %>%
  ggplot(aes(x = tissue, y = v_gene, fill = proportion)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "#FAEBEF", high = "#84223C") +
  geom_text(aes(label = round(proportion, 1)), color = "black", size = 5, fontface = "bold") +
  labs(x = "Timepoint",
       y = "V Gene",
       fill = "% of Total") +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 20), legend.title = element_text(size = 15, face = "bold", color = "black"))

#save
ggsave(filename = paste0(figures.dir, "top5_IGHV_by_timepoint_heatmap.svg"), width = 10, height = 6)
```

Top 10 Frequency
```{r}
top10_summary <- data %>%
  count(tissue, v_gene) %>%
  group_by(tissue) %>%
  mutate(proportion = n / sum(n) * 100) %>%
  top_n(10, n) %>%
  summarise(
    top10_pct    = sum(proportion),      # total % accounted for by top 10
    n_vgenes     = n(),                  # should be 10 for each tissue
    .groups      = "drop"
  )

print(top10_summary)


top10_summary <- data %>%
  count(tissue, v_gene) %>%
  group_by(tissue) %>%
  mutate(proportion = n / sum(n) * 100) %>%
  mutate(total_unique_vgenes = n_distinct(v_gene)) %>%   # total unique v genes per tissue
  top_n(10, n) %>%
  summarise(
    top10_pct          = round(sum(proportion), 1),
    total_unique_vgenes = first(total_unique_vgenes),     # same value for all rows so just take first
    n_vgenes           = n(),
    .groups            = "drop"
  )

print(top10_summary)
```
normalised trajectory
```{r}
v_genes_of_interest <- c("IGHV3-88", "IGHV3-NL_17", "IGHV3-50", "IGHV3-NL_11", "IGHV4-NL_33", 
                         "IGHV7-86", "IGHV4-149", "IGHV3-172", "IGHV4-117", "IGHV3-NL_19")

vgene_normalized <- data %>%
  filter(v_gene %in% v_genes_of_interest) %>%
  group_by(animal_ID, tissue, v_gene) %>%
  summarise(n = n(), .groups = "drop") %>%
  left_join(
    data %>%
      group_by(animal_ID, tissue) %>%
      summarise(total = n(), .groups = "drop"),
    by = c("animal_ID", "tissue")
  ) %>%
  mutate(freq = n / total * 100)

vgene_summary <- vgene_normalized %>%
  group_by(tissue, v_gene) %>%
  summarise(
    mean_freq = mean(freq, na.rm = TRUE),
    se_freq   = sd(freq, na.rm = TRUE) / sqrt(n()),
    .groups   = "drop"
  )

ggplot(vgene_summary, aes(x = tissue, y = mean_freq,
                           color = v_gene, group = v_gene)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_freq - se_freq,
                    ymax = mean_freq + se_freq),
                width = 0.2) +
  scale_color_observable() +
  labs(x     = "Timepoint",
       y     = "Mean V Gene Frequency (%)",
       color = "V Gene") +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 20), legend.text  = element_text(size = 14))

#save
ggsave(filename = paste0(figures.dir, "top_5_D16_IGHV_trajectory_norm_3.svg"),  width = 10, height = 6)
```

IGHV richness
```{r}
data %>%
  group_by(tissue) %>%
  summarise(n_unique_vgenes = n_distinct(v_gene)) %>%
  ggplot(aes(x = tissue, y = n_unique_vgenes, fill = tissue)) +
  geom_col() +
  geom_text(aes(label = n_unique_vgenes),    # adds number on top of each bar
            vjust = -0.5, size = 5) +
  scale_fill_malaria() +
  labs(x = NULL,
       y = "Number of Unique V Genes") +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18), legend.position = "none")

#save
ggsave(filename = paste0(figures.dir, "V_gene_richness_time.svg"),  width =6 , height = 6)

```

******************************************
Clonotype analysis - scSanger (only looking at D16 and D49)
-In order to run diversity analysis/normalization, need to drop animals that do not have enough sequences/have many unique clonotypes.
-Therefore, looking at number of sequences per animal and number of unique clonotypes is important
```{r}
complete_animals <- c("1010", "2008", "5013", "6017", "6003", "9002")

#sequence per animal
sc_summary <- data_sc %>%
  filter(animal_ID %in% complete_animals) %>%
  group_by(animal_ID, timepoint, clone_id) %>%
  summarise(clone_size = n(), .groups = "drop")

sc_summary %>%
  arrange(animal_ID, clone_size) %>%           # sort smallest to largest within each animal
  mutate(clone_id = factor(clone_id,           # fix order by converting to factor
                           levels = unique(clone_id))) %>%
  ggplot(aes(x = animal_ID, y = clone_size, fill = clone_size)) +
  geom_bar(position="stack", stat="identity", color = "black") +
  scale_fill_viridis_c() +
  labs(x = "Animal ID", y = "# sequences", fill = "Clonotype size") +
  stat_summary(fun = sum, aes(label = after_stat(y), group = animal_ID), geom = "text", vjust = -.2) +
  theme_classic() +
  theme(axis.text = element_text(size = 16), axis.title = element_text(size = 18), strip.text = element_text(size = 18), legend.title=element_text(size=14), legend.text = element_text(size = 14))+
  facet_wrap(~timepoint, scales = "free_x")

#save
ggsave(filename = paste0(figures.dir, "sc_clonotype_variability.svg"), width = 12, height = 7)
```

Clonotype analysis - Bulk
```{r}
#sequence per animal
bulk_summary <- data_bulk %>%
  filter(!animal_ID=="6005")%>%
  group_by(animal_ID, tissue, clone_id) %>%
  summarise(clone_size = n(), .groups = "drop")

bulk_summary %>%
  arrange(animal_ID, clone_size) %>%           # sort smallest to largest within each animal
  mutate(clone_id = factor(clone_id,           # fix order by converting to factor
                           levels = unique(clone_id))) %>%
  ggplot(aes(x = animal_ID, y = clone_size, fill = clone_size)) +
  geom_bar(position="stack", stat="identity", color = "black") +
  scale_fill_viridis_c() +
  labs(x = "Animal ID", y = "# sequences", fill = "Clonotype size") +
  stat_summary(fun.y = sum, aes(label = after_stat(y), group = animal_ID), geom = "text", vjust = -.2) +
  theme_classic() +
  facet_wrap(~tissue, scales = "free_x") +
   theme(axis.text = element_text(size = 16), axis.title = element_text(size = 18), strip.text = element_text(size = 18), legend.title=element_text(size=14), legend.text = element_text(size = 14))

#save
ggsave(filename = paste0(figures.dir, "bulk_clonotype_variability.svg"), width = 16, height = 8)


```



Clonal Diversity
Rodrigo's code: https://rodrigarc.github.io/rabies_mRNA/results/lab_book/bcr-repertoire-analysis.html

Chao1 - Bulk
```{r}

bulk_summary <- bulk_summary %>%
  mutate(tissue = str_remove(tissue, "_D66"))%>%
  mutate(ID = paste0(animal_ID, "_", tissue))

# Run chao1 100 times and collect the values
chaox100 <- function(x){
  replicate(100, {
  subsample <-  rrarefy(x,min(rowSums(x)))
  chao <- estimateR(subsample)
  return(chao)
  })}

# Extract columns from df, remove NAs, and reorder based on decreasing value
extract_and_order <- function(df, column) {
  group <- df[[column]][!is.na(df[[column]])]
  group <- group[order(group, decreasing = TRUE)]
  return(group)
}

# Save as list the clonal size for each individual
list_clonal_size <- list()
for(i in unique(bulk_summary$ID)){
  list_clonal_size[[i]] <- bulk_summary %>% filter(ID == i) %>% pull(clone_size)
}

# Remove individuals with less than 20 sequences since that sample size is too small, arbitrary number based on our sequencing depth
  {
    filt_ls_clonal_size <- list_clonal_size[lapply(list_clonal_size, function(x) sum(x)) > 20]
  filt_ls_clonal_size_chao100x <- lapply(filt_ls_clonal_size, function(x) x[order(x, decreasing = TRUE)])
  max_length <- do.call(max, lapply(filt_ls_clonal_size_chao100x, length))
  
  for(l in names(filt_ls_clonal_size_chao100x)){
    length(filt_ls_clonal_size_chao100x[[l]]) <- max_length
  }
  list_names <- names(filt_ls_clonal_size_chao100x)
  filt_ls_clonal_size_chao100x <- lapply(filt_ls_clonal_size_chao100x, as.data.frame)
  filt_ls_clonal_size_chao100x <- do.call(cbind, filt_ls_clonal_size_chao100x)
  colnames(filt_ls_clonal_size_chao100x) <- list_names
  input_chao <- as.data.frame(t(filt_ls_clonal_size_chao100x))
  input_chao[is.na(input_chao)] <- 0
}
```

iNEXT
```{r}
inextx100 <- function(x, level, subsampling){
  inext_D <- list()
  replicate(100, {
  subsample <- rrarefy(unlist(x), subsampling)
  subsample <-  subsample[order(subsample, decreasing = T)]
  inext_D[[i]] <- estimateD(subsample, q = c(0,1,2), datatype = "abundance", base = "size", level = level, nboot = 0)
  return(inext_D)})
}

inext_obj <- lapply(filt_ls_clonal_size, inextx100, level = NULL, subsampling = 24)
inext_obj <- lapply(inext_obj, data.table::rbindlist)
inext_rep_100 <- data.table::rbindlist(inext_obj, idcol = TRUE, fill = TRUE) %>%
  mutate(ID = .id, Group =plyr::mapvalues(ID, from = c("1010_BM", "1010_blood", "2001_BM", "2001_blood", "2008_BM", "2008_blood", "2010_BM", "2010_blood", "5009_BM", "5009_blood", "5013_BM", "5013_blood", "6003_BM", "6003_blood", "6005_BM", "6005_blood", "6017_BM", "6017_blood", "7010_BM", "7010_blood", "8002_BM", "9002_BM", "9002_blood"), to = c("BM", "blood", "BM", "blood", "BM", "blood", "BM", "blood", "BM", "blood", "BM", "blood", "BM", "blood", "BM", "blood", "BM", "blood", "BM", "blood", "BM","BM", "blood"))) %>%
  group_by(ID, Order.q,Group) %>%
  summarise(avg_estimator = mean(qD))

inext_rep_100 %>%
  write.csv(paste0(result.dir,"diversity_per_bulk.csv"), row.names = FALSE)

inext_rep_100 %>%
  mutate(Order.q = plyr::mapvalues(Order.q, from = c(0,1,2), to = c("Clonotype richness", "Shannon diversity", "Simpson diversity"))) %>%
  ggplot(aes(Group, avg_estimator, group = Group, fill = Group)) + #issue bc not into groups; how to make by timepoint
  geom_boxplot(outlier.shape = NA, aes(middle = mean(avg_estimator))) +
  geom_point(position = position_jitter(.2), size = 3) +
 stat_compare_means(method = "wilcox.test", label = "p.signif", comparisons = list(c("BM", "blood")), paired = FALSE, label.y = c(20)) +
  theme_classic() +
  labs(y = "Estimator\n(100x mean)", x = "") +
  theme(legend.position = "none",axis.text = element_text(size = 16), axis.title = element_text(size = 18), strip.text = element_text(size = 18)) +
  facet_wrap(~Order.q)
 
#save
ggsave(filename = paste0(figures.dir, "bulk_clonotype_diversity.svg"), width = 10, height = 6)


```



Chao1 - Single-cell
```{r}
sc_summary <- sc_summary %>%
  mutate(ID = paste0(animal_ID, "_", timepoint))%>%
  filter(!animal_ID=="6017")

# Run chao1 100 times and collect the values
chaox100 <- function(x){
  replicate(100, {
  subsample <-  rrarefy(x,min(rowSums(x)))
  chao <- estimateR(subsample)
  return(chao)
  })}

# Extract columns from df, remove NAs, and reorder based on decreasing value
extract_and_order <- function(df, column) {
  group <- df[[column]][!is.na(df[[column]])]
  group <- group[order(group, decreasing = TRUE)]
  return(group)
}

# Save as list the clonal size for each individual
list_clonal_size <- list()
for(i in unique(sc_summary$ID)){
  list_clonal_size[[i]] <- sc_summary %>% filter(ID == i) %>% pull(clone_size)
}

# Remove individuals with less than 20 sequences since that sample size is too small, arbitrary number based on our sequencing depth
  {
    filt_ls_clonal_size <- list_clonal_size[lapply(list_clonal_size, function(x) sum(x)) > 20]
  filt_ls_clonal_size_chao100x <- lapply(filt_ls_clonal_size, function(x) x[order(x, decreasing = TRUE)])
  max_length <- do.call(max, lapply(filt_ls_clonal_size_chao100x, length))
  
  for(l in names(filt_ls_clonal_size_chao100x)){
    length(filt_ls_clonal_size_chao100x[[l]]) <- max_length
  }
  list_names <- names(filt_ls_clonal_size_chao100x)
  filt_ls_clonal_size_chao100x <- lapply(filt_ls_clonal_size_chao100x, as.data.frame)
  filt_ls_clonal_size_chao100x <- do.call(cbind, filt_ls_clonal_size_chao100x)
  colnames(filt_ls_clonal_size_chao100x) <- list_names
  input_chao <- as.data.frame(t(filt_ls_clonal_size_chao100x))
  input_chao[is.na(input_chao)] <- 0
}
```

iNEXT
```{r}
inextx100 <- function(x, level, subsampling){
  inext_D <- list()
  replicate(100, {
  subsample <- rrarefy(unlist(x), subsampling)
  subsample <-  subsample[order(subsample, decreasing = T)]
  inext_D[[i]] <- estimateD(subsample, q = c(0,1,2), datatype = "abundance", base = "size", level = level, nboot = 0)
  return(inext_D)})
}

inext_obj <- lapply(filt_ls_clonal_size, inextx100, level = NULL, subsampling = 24)
inext_obj <- lapply(inext_obj, data.table::rbindlist)
inext_rep_100 <- data.table::rbindlist(inext_obj, idcol = TRUE, fill = TRUE) %>%
  mutate(ID = .id, Group =plyr::mapvalues(ID, from = c("1010_D16", "1010_D49", "2008_D16", "2008_D49", "5013_D16", "5013_D49", "6003_D16", "6003_D49", "6017_D16", "6017_D49", "9002_D16", "9002_D49"), to = c("D16", "D49", "D16", "D49", "D16", "D49", "D16", "D49", "D16", "D49", "D16", "D49"))) %>%
  group_by(ID, Order.q,Group) %>%
  summarise(avg_estimator = mean(qD))

inext_rep_100 %>%
  write.csv(paste0(result.dir,"diversity_per_SC.csv"), row.names = FALSE)

inext_rep_100 %>%
  mutate(Order.q = plyr::mapvalues(Order.q, from = c(0,1,2), to = c("Clonotype richness", "Shannon diversity", "Simpson diversity"))) %>%
  ggplot(aes(Group, avg_estimator, group = Group, fill = Group)) + #issue bc not into groups; how to make by timepoint
  geom_boxplot(outlier.shape = NA, aes(middle = mean(avg_estimator))) +
  geom_point(position = position_jitter(.2), size = 3) +
 stat_compare_means(method = "wilcox.test", label = "p.signif", comparisons = list(c("D16", "D49")), paired = FALSE, label.y = c(50)) +
  theme_classic() +
  scale_fill_bmj()+
  labs(y = "Estimator\n(100x mean)", x = "") +
  theme(legend.position = "none", axis.text = element_text(size = 16), axis.title = element_text(size = 18), strip.text = element_text(size = 18)) +
  facet_wrap(~Order.q)

#save
ggsave(filename = paste0(figures.dir, "sc_clonotype_diversity.svg"), width = 10, height = 6)


```

mAbs comparison -> alignment of single-cell
```{r}
mAbs<- c("5013_03_E08","6017_P2_G11", "5013_02_H01")
daa <- data %>%
  mutate(aa_pos52 = str_sub(VDJ_aa, 52, 52))%>%
  filter(sequence_id %in% mAbs)
```


```{r}
sessionInfo()
```